### 2015-04-07 contract-language.tex
Formatting edits
Bug fix in proof.

\begin{figure}
\vspace{-7pt}
\begin{smathpar}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

Note that the axiomatic definition of an execution ($\E$) provides
interpretations for primitive relations (eg: $\visZ$) that occur free in
contract formulas, and also fixes the domain of quantification to set of all
effects ($\EffSoup$) observed during the program execution. As such, $\E$ is a
potential model for any first-order formula ($\cv$) expressible in our contract
language. If $\E$ is indeed a valid model for $\cv$ (written as $\E \models
\cv$), we say that the execution $\E$ satisfied the contract $\cv$:
\begin{definition}
An axiomatic execution $\E$  satisfies a contract $\cv$ if and
only if $E \models \cv$.
\end{definition}

@@

\begin{itemize}[noitemsep]

\item \textbf{Eventual consistency}: Eventually consistent operations can
	be satisfied as long as the client can reach at least one replica. In the
	bank account example, \cf{deposit} is an eventually consistent operation.
	While an \ecds typically offer \emph{basic} eventual consistency with all
	possible anomalies, we assume that our store provides stronger semantics that
	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
	\emph{causal cut} of the updates. This semantics can be formally captured in
	terms of the following contract definition:

@@

\subsection{Contract Classification}
\label{sec:classify}

\newcommand{\DDe}[1]{#1}
\begin{figure}
\vspace{-10pt}
\begin{smathpar}
\begin{array}{cc}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  &

\RuleTwo
{\DDe{\cv} \le \DDe{\ecc}}
{{\sf EventuallyConsistent}(\cv)} \\ \\

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ecc}
\quad \DDe{\cv} \le \DDe{\ccc}}
{{\sf CausallyConsistent}(\cv)} &

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\vspace{-3mm}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{definition}
An axiomatic execution $\small \E = \Exec$ is well-formed if the
following axioms ($\Delta$) hold:

\begin{itemize}[noitemsep]
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

\begin{proof}
  Hypothesis:
  \begin{smathpar}
  \begin{array}{cr}
    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E', \sigma \pll \Sigma} & H\npp\\
    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
  \end{array}
  \end{smathpar}
  Since $\tau$ is the contract class of $\cv$, by inversion, we have
  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
  \end{array}
  \end{smathpar}
   Since $\eff$ denotes new effect, it is a fresh variable that does
   not occur free in $\Delta$. From $H3$, after instantiating bound
   $\cureff$ with $\eff$, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
      & H\npp\\
  \end{array}
  \end{smathpar}
  Due to the soundness of natural deduction for first-order logic,
  $H4$ implies that for all models $\mathcal{M}$ such that
  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \E' \models \Delta & H\npp\\
  \end{array}
  \end{smathpar}
  Proof follows from $H1$, $H5$, and $H4$.
  \hfill \qed
\end{proof}

>>>>

\begin{figure}
\begin{smathpar}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

Note that the axiomatic definition of an execution ($\E$) provides
interpretations for primitive relations (eg: $\visZ$) that occur free in
contract formulas, and also fixes the domain of quantification to set of all
effects ($\EffSoup$) observed during the program execution. As such, $\E$ is a
potential model for any first-order formula ($\cv$) expressible in our contract
language. If $\E$ is indeed a valid model for $\cv$ (expressed using the
model-theoretic consequence relation as $\E \models \cv$), we say that the
execution $\E$ satisfied the contract $\cv$:
\begin{definition}
An axiomatic execution $\E$ is said to satisfy a contract $\cv$ if and
only if $E \models \cv$.
\end{definition}

@@

\begin{itemize}
\setlength{\itemsep}{2pt}

\item \textbf{Eventual consistency}: Eventually consistent operations can
	be satisfied as long as the client can reach at least one replica. In the
	bank account example, \cf{deposit} is an eventually consistent operation.
	While eventually consistent data stores typically offer \emph{basic} eventual
	consistency with all possible anomalies, we assume that our store provides
	stronger semantics that remain highly-available~\cite{BailisHAT,COPS}; the
	store always exposes a \emph{causal cut} of the updates. This semantics can
	be formally captured in terms of the following contract definition:

@@

\subsection{Contract Comparison and Classification}

\newcommand{\DDe}[1]{#1}
\begin{figure}
\begin{smathpar}
\begin{array}{c}
\vspace{3mm}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  \qquad

\RuleTwo
{\DDe{\cv} \le \DDe{\ecc}}
{{\sf EventuallyConsistent}(\cv)} \\

\vspace{3mm}
\RuleTwo
{\DDe{\cv} \not\le \DDe{\ecc}
\quad \DDe{\cv} \le \DDe{\ccc}}
{{\sf CausallyConsistent}(\cv)} \qquad

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\vspace{-5mm}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{definition}
An axiomatic executions $\small \E = \Exec$ is said to be well-formed if the
following axioms ($\Delta$) hold:

\begin{itemize}
\setlength{\itemsep}{2pt}
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

\begin{proof}
  Hypothesis:
  \begin{smathpar}
  \begin{array}{cr}
    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E', \sigma \pll \Sigma} & H\npp\\
    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
  \end{array}
  \end{smathpar}
  Since $\tau$ is the contract class of $\cv$, by inversion, we have
  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
  \end{array}
  \end{smathpar}
   Since $\eff$ denotes new effect, it is a fresh variable that does
   not occur free in $\Delta$. From $H3$, after instantiating bound
   $\cureff$ with $\eff$, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
      & H\npp\\
  \end{array}
  \end{smathpar}
  Due to the soundness of natural deduction for first-order logic,
  $H4$ implies that for all models $\mathcal{M}$ such that
  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \E' \models \Delta & H\npp\\
  \end{array}
  \end{smathpar}
  Proof follows from $H1$, $H4$, and $H3$.
  \qed
\end{proof}
